---
title: "MANORM Annotation Report"
output: html_document

params:
  peak_in: ""
  PeakIdnt: ""
  samplename: ""
  background: ""
  pval: .005
  FC: 1
  incd_rRNA: T
---
<!-- # Load libraries, set theme   -->
```{r environment, include=F ,echo=F,warning=F,message=FALSE}
# rm(list=setdiff(ls(), "params"))
suppressMessages(library(rmarkdown))
suppressMessages(library(stringr))
#suppressMessages(library(rtracklayer))
#suppressMessages(library(VariantAnnotation,quietly = T,verbose = F,warn.conflicts = F,logical.return = F))
suppressMessages(library(rmarkdown))
suppressMessages(library(ggplot2,quietly = T,verbose = F,warn.conflicts = F,logical.return = F))
suppressMessages(library("viridis",quietly = T,verbose = F,warn.conflicts = F,logical.return = F))
#suppressMessages(library(edgeR,quietly = T,verbose = F))
#suppressMessages(library('GenomicFeatures',quietly = T,verbose = F,warn.conflicts = F,logical.return = F))
#suppressMessages(library('rtracklayer',quietly = T,verbose = F,warn.conflicts = F,logical.return = F))
#suppressMessages(library(matrixStats,quietly = T,verbose = F,warn.conflicts = F,logical.return = F))
suppressMessages(library(plyr,quietly = T,verbose = F,warn.conflicts = F,logical.return = F))
suppressMessages(library(tidyr,quietly = T,verbose = F,warn.conflicts = F,logical.return = F))
#suppressMessages(library(fitdistrplus,quietly = T,verbose = F,warn.conflicts = F,logical.return = F))
suppressMessages(library(stringr,quietly = T,verbose = F,warn.conflicts = F,logical.return = F))
suppressMessages(library(data.table))
#suppressMessages(library(reshape))
suppressMessages(library(knitr))
suppressMessages(library(stringi))
#suppressMessages(library(BSgenome))
#suppressMessages(library(biomaRt))
suppressMessages(library(plotly))
suppressMessages(library(tidyr))
#suppressMessages(library(GenomicRanges))
suppressMessages(library(RColorBrewer))
suppressMessages(library('gplots'))
suppressMessages(library(ggpubr))
suppressMessages(library(circlize))
#suppressMessages(library('regioneR'))
suppressMessages(library(scales))
suppressMessages(library(dplyr))

blank_theme <- theme_minimal()+
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.grid=element_blank(),
  axis.ticks = element_blank(),
  plot.title=element_text(size=14, face="bold")
  )
```

<!-- # Set params, output   -->

```{r echo=F, message=FALSE, warning=FALSE, eval=T, include=F,results=F,}
peak_in=read.table(params$peak_in, sep = '\t', header=TRUE, fill=TRUE)
samplename=params$samplename
background=params$background
pval_Cut=as.numeric(params$pval)
FC_cut=as.numeric(params$FC)
incd_rRNA=params$incd_rRNA
```

```{r echo=F, message=FALSE, warning=FALSE, eval=T, include=F,results=F,}
nc_color= c('yRNA'= "#A6CEE3",
  'snRNA'= "#1F78B4",
  'snoRNA'="#B2DF8A" ,
  'srpRNA'= "#33A02C",
  'tRNA'=   "#FB9A99",
  '7SKRNA'= "#E31A1C",
  'scRNA'=  "#FDBF6F",
  'other_scRNA'=  "#FDBF6F",
  'sncRNA'= "#FF7F00",
  'scaRNA'= "#CAB2D6",
  'miRNA'= "#6A3D9A" ,
  'vaultRNA'= "#FFFF99",
  'sRNA'= "#B15928"   ,
  'rRNA'=  "#1fb487"  ,
  'miscRNA'= "grey",
  'misc_RNA'= "grey")

class_color=c(
   "ncRNA"="#1B9E77",
   "rRNA" ="#1fb487",
   "yRNA"= "#2bdba6" ,
   "protein_coding: Exonic"="#D95F02",
   "Repeat Element"="#7570B3" ,
   "pseudogene"="#E7298A", 
   "Antisense Feature"="#66A61E", 
   "protein_coding: Intronic"="#E6AB02", 
   "lncRNA"="#A6761D", 
   "no Feature"="#666666")
```

```{r echo=F, message=FALSE, warning=FALSE, eval=T, include=T, results=F,}
# Identification of CLIP peaks: `r toString(params$samplename)`    

Peaks_Anno = peak_in

Peaks_Anno$Same_Comb_type = 
    factor(peak_in$Same_Comb_type, 
           levels = c("rRNA", "ncRNA", "protein_coding: Exonic", "Repeat Element","pseudogene",
                      "Antisense Feature", "protein_coding: Intronic", "lncRNA: Intronic", "lncRNA: Exonic","no Feature"))  

Peaks_Anno$Oppo_Comb_type = 
    factor(peak_in$Oppo_Comb_type, 
           levels = c("rRNA", "ncRNA", "protein_coding: Exonic", "Repeat Element","pseudogene",
                      "Antisense Feature", "protein_coding: Intronic", "lncRNA: Intronic", "lncRNA: Exonic","no Feature"))  

###### remove chrM clip peaks
Peaks_Anno=Peaks_Anno[Peaks_Anno$chrom%in%'chrM'==F,] 


#### Update ncRNA categories

Peaks_Anno$Same_ncRNA[grep("HY", Peaks_Anno$Same_ncRNA)] <- 'yRNA'
Peaks_Anno$Same_ncRNA[grep("tRNA", Peaks_Anno$Same_ncRNA)] <- 'tRNA'
Peaks_Anno$Same_ncRNA[grep("7SK", Peaks_Anno$Same_ncRNA)] <- '7SKRNA'
Peaks_Anno$Same_ncRNA <-  gsub('miRNA,','',Peaks_Anno$Same_ncRNA)
Peaks_Anno$Same_ncRNA <-  gsub(' miRNA','',Peaks_Anno$Same_ncRNA)
Peaks_Anno$Same_ncRNA <-  gsub('misc_RNA,','',Peaks_Anno$Same_ncRNA)
Peaks_Anno$Same_ncRNA <-  gsub(' misc_RNA','',Peaks_Anno$Same_ncRNA)
Peaks_Anno$Same_ncRNA <-  gsub(' ','',Peaks_Anno$Same_ncRNA)
Peaks_Anno$Same_ncRNA <-  gsub(',','',Peaks_Anno$Same_ncRNA)

#Peaks_Anno$Same_ncRNA[grep("BK000964", Peaks_Anno$chrom)] <- 'rRNA'
Peaks_Anno=Peaks_Anno[!Peaks_Anno$chrom%in%'GL',]
Peaks_Anno=Peaks_Anno[!Peaks_Anno$chrom%in%'JH',]
Peaks_Anno=Peaks_Anno[!Peaks_Anno$chrom%in%'K',]


Peaks_Anno$PlotColumn=Peaks_Anno$Same_Comb_type
Peaks_Anno$PlotColumn=as.character(Peaks_Anno$PlotColumn)
Peaks_Anno[Peaks_Anno$Same_Repeat%in%'rRNA','PlotColumn']='rRNA'
Peaks_Anno[Peaks_Anno$Same_ncRNA%in%'yRNA','PlotColumn']='yRNA'

Peaks_Anno$Counts_Unique <- Peaks_Anno$Counts_unique_Sample
Peaks_Anno$Counts_fracMM <- Peaks_Anno$Counts_fracMM_Sample
Peaks_Anno$Counts_Total <- Peaks_Anno$Counts_fracMM_Sample_Total
```


<!-- ## Processing data    -->
# Differential CLIP peaks identified by MAnorm (`r samplename`-`r background`)  

MAnorm used to identify Peaks preferentially expressed in the `r (samplename)` samples compared to `r (background)`.       

## I. Identify preferentialy expressed CLIP peaks  
```{r ,fig.asp=.5, fig.align='center',fig.height=3.5, fig.width=7,echo=F,warning=F,eval=T,include=T}





g1=ggplot(Peaks_Anno,aes(x=m_value,y=-log10(p_value),color=`PlotColumn`)) +
  geom_point(size=0.5)+
  theme_classic()+
  xlab(paste0("log2 M-value \n (",samplename,"-",background,")"))+
  ylab(paste0("-log10 P-Value"))+
  labs(color='RNA Type')+
  geom_hline(yintercept=-log10(pval_Cut),size=.5,col='black',linetype="dashed")+
  geom_vline(xintercept=(FC_cut),size=.5,col='black',linetype="dashed")+
  geom_vline(xintercept=(-FC_cut),size=.5,col='black',linetype="dashed")+
  scale_color_manual(values=class_color)

g2=ggplot(Peaks_Anno,aes(x=log2(Peaks_Anno[,'normalized_read_density_sample']),y=m_value,color=-log10(p_value)))+
  geom_point(size=0.5)+
  theme_classic()+
  xlab(paste0("log2 normalized Counts \n (",samplename,")"))+
  ylab(paste0("log2 M-value \n(",samplename,"-",background,")"))+
  scale_colour_gradient(  low = ("blue"),  high = ("red"),
                           na.value = "grey50",guide = "colourbar")+
  geom_hline(yintercept=(FC_cut),size=.5,col='black',linetype="dashed")+
  geom_hline(yintercept=(-FC_cut),size=.5,col='black',linetype="dashed")
    
g3=ggplot(Peaks_Anno,aes(x=log2(Peaks_Anno[,'normalized_read_density_sample']),y=m_value,color=`PlotColumn`))+
  geom_point(size=0.5)+
  theme_classic()+
  xlab(paste0("log2 normalized Counts \n (",samplename,")"))+
  ylab(paste0("log2 M-value \n(",samplename,"-",background,")"))+
  labs(color='RNA Type')+
  geom_hline(yintercept=(FC_cut),size=.5,col='black',linetype="dashed")+
  geom_hline(yintercept=(-FC_cut),size=.5,col='black',linetype="dashed")+
  scale_color_manual(values=class_color) 
if (nrow(Peaks_Anno)<5000) {
  g1=ggplotly(g1,tooltip = c("text"))
  g2=ggplotly(g2,tooltip = c("text"))
  g3=ggplotly(g3,tooltip = c("text"))
  }
g1;g2;g3
```

```{r ,fig.asp=.5, fig.align='center',echo=F,warning=F,eval=T,include=T}
MPeak=median(Peaks_Anno$end - Peaks_Anno$start)
```

```{r fig.asp=.1, fig.align='center',fig.height=.1,echo=F,warning=F,eval=T,include=T}

MPeakPrint=paste0("Median Peak width: ",MPeak,"nt\n")

if (MPeak>=50&MPeak<=100) {
matrix(1:9, nrow = 3, ncol = 3)%>%as.data.frame()%>%ggplot(.,aes(x=V1,y=V2))+geom_text(x=.4, y=.1, label=MPeakPrint,size=4.5)+theme_void()
}
```

```{r fig.asp=.5, fig.align='center', fig.width=5,echo=F,warning=F,eval=T,include=T}
if (MPeak<50|MPeak>100) {
  
txt1=paste0( "Consider modifying MAnorm values\n",
                             "-w ",MPeak," and -d ",MPeak/2
                                                          )
txt2=paste0(
                             "Default Values:\n",
                             "-w 50 and -d 25"
                             )
  
matrix(1:3, nrow = 3, ncol = 3)%>%as.data.frame()%>%ggplot(.,aes(x=V1,y=V2))+
  geom_text(x=.4, y=.8, label=MPeakPrint,size=6,color='red')+
  geom_text(x=.4, y=.55, label=txt1,size=6,color='red')+
  geom_text(x=.4, y=.2, label=txt2,size=5,color='black')+
  theme_void()

}

```

<br>  




```{r ,fig.asp=.5, fig.align='center',echo=F,warning=F,eval=T,include=T}
Peaks_Anno=Peaks_Anno[Peaks_Anno$p_value<pval_Cut&abs(Peaks_Anno$m_value)>FC_cut,]
```




## Peaks with a P-value < `r pval_Cut` and |log~2~(fold change)| > `r FC_cut` were selected, resulting in `r toString(nrow(Peaks_Anno))` peaks.   

<br>

### II. Overview of targets by CLIP peak category   


```{r PeakCategories, echo=F, include=T, eval=T, fig.align='default', fig.ncol=2, fig.height=3.5, fig.width=7, out.height=200, out.width="50%", fig.show="hold", layout='l-screen', results = FALSE, message=FALSE , warning=F}

print(head(Peaks_Anno))

##Plot 1
#if (params$PeakIdnt=='UNIQUE'){
#  x1= ggplot(Peaks_Anno,
#            aes(x=Counts_Unique, 
#                y=Length,
#                color=PlotColumn,
#                text=Same_external_gene_name)) +
#      geom_point(size=1) +
#      # scale_x_continuous(trans = "log2",limits = c(.5,NA))+
#      scale_y_continuous(trans = "log2",
#                          limits = c(NA,NA),
#                          oob = squish_infinite) +
#      theme_classic() +
#      ylab("Peak Width") +
#      xlab("Peak Counts (unique)") +
#      ggtitle("Comparison of peak counts and Peak Width") +
#      scale_color_manual(values=class_color) 

#} else if (params$PeakIdnt=='ALL'){
#  x1= ggplot(Peaks_Anno,
#             aes(x=Counts_fracMM,
#                 y=Length,
#                 color=PlotColumn,
#                 text=Same_external_gene_name)) +
#      geom_point(size=1)+
#      # scale_x_continuous(trans = "log2",limits = c(4,NA))+
#      scale_y_continuous(trans = "log2",
#                          limits = c(NA,NA),
#                          oob = squish_infinite) + 
#      theme_classic() +
#      ylab("Peak Width") +
#      xlab("Peak Counts (Unique+ FracMM)") +
#      ggtitle("Comparison of peak counts and Peak Width") +
#      labs(color='RNA Type') +
#      scale_color_manual(values=class_color) 
#}

##Plot 2
#x2= ggplot(Peaks_Anno,
#           aes(x=Length)) +
#    geom_histogram(bins = 100) +
#    # scale_x_continuous(trans = "log2",limits = c(20,NA)) + 
#    theme_classic() +
#    ylab("# of peaks") +
#    xlab("Peak Width") +
#    ggtitle("Distribution of peak Width")

##Plot 3
#x3= ggplot(Peaks_Anno,
#           aes(x=Length,
#               line=PlotColumn,
#               color=PlotColumn)) +
#    geom_density(size=.5) +
#    # scale_x_continuous(trans = "log2",limits = c(8,NA)) + 
#    theme_classic() +
#    ylab("Density") +
#    xlab("Peak Width") +
#    ggtitle("Distribution of peak Width") +
#    labs(color='RNA Type') +
#    scale_color_manual(values=class_color) 

##Plot 4
#x4= ggplot(Peaks_Anno,
#           aes(x=Length,
#               line=PlotColumn,
#               color=PlotColumn)) +
#    geom_freqpoly(bins = 100)+
#    # scale_x_continuous(trans = "log2",limits = c(8,NA)) + 
#    theme_classic() +
#    ylab("# of peaks") +
#    xlab("Peak Width") +
#    ggtitle("Distribution of peak Width") +
#    labs(color='RNA Type') +
#    scale_color_manual(values=class_color) 

##Plot 5
#if (params$PeakIdnt=='UNIQUE'){
#  x5= ggplot(Peaks_Anno,
#            aes(x=Counts_Unique,
#                line=PlotColumn,
#                color=PlotColumn)) +
#      geom_density() +
#      # scale_x_continuous(trans = "log2",limits = c(.25,NA)) + 
#      theme_classic() +
#      ylab("Density") +
#      xlab("Peak Counts (unique)") +
#      ggtitle("Distribution of peak counts") + 
#      labs(color='RNA Type') +
#      scale_color_manual(values=class_color)

#} else if (params$PeakIdnt=='ALL'){
#  x5= ggplot(Peaks_Anno,
#             aes(x=Counts_fracMM,
#                 line=PlotColumn,
#                 color=PlotColumn)) +
#      geom_density() +
#      # scale_x_continuous(trans = "log2",limits = c(.25,NA)) + 
#      theme_classic() +
#      ylab("Density") +
#      xlab("Peak Counts (Unique+ FracMM)") +
#      ggtitle("Distribution of peak counts") +
#      labs(color='RNA Type') +
#      scale_color_manual(values=class_color) 
#}

##Plot all graphs
#if (nrow(Peaks_Anno)>1) {
#  x1=x1+scale_x_continuous(trans = "log2",limits = c(4,NA));  
#  x2=x2+scale_x_continuous(trans = "log2",limits = c(20,NA));  
#  x3=x3+scale_x_continuous(trans = "log2",limits = c(8,NA));
#  x4=x4+scale_x_continuous(trans = "log2",limits = c(8,NA));  
#  x5=x5+scale_x_continuous(trans = "log2",limits = c(.25,NA))
#}
#  x1;x2;x3;x4;x5
```


### III.  Number of peaks assigned to each category  


**III.1 Peak counts do include rRNA peaks**   

```{r PeaksPerCategory, fig.asp=.5, fig.align='center',fig.height=4, fig.width=8,echo=F,warning=F,eval=T,include=T}
p=Peaks_Anno
p$PlotColumn=p$Same_Comb_type
if (incd_rRNA==FALSE) {
  p=p[!p$Same_Comb_type%in%'rRNA',]
}
if (nrow(p)>0) {
#change types
p[is.na(p$PlotColumn),'PlotColumn']='no Feature'
colnames(p)[colnames(p)%in%'PlotColumn']='RNA_Type'
p$RNA=''
#plot
gg=ggplot(p,aes(fill=RNA_Type,x=RNA) ) +
  geom_bar(position="stack",width=.5) +
  theme_classic()+ggtitle("BioType: # of peaks per category") +
  scale_fill_manual(values=class_color) 
ggplotly(gg)
}else{print("No ncRNA")}
```


```{r PeaksPer_NCRNA_Category, fig.asp=.5, fig.align='center',fig.height=4, fig.width=8,echo=F,warning=F,eval=T,include=T}
p=Peaks_Anno
p$PlotColumn=p$Same_Comb_type

###phil need else statement?###
#check if this makes sense?
if (incd_rRNA==FALSE) {
  p=p[!p$Same_Comb_type%in%'rRNA',]
} else{
  p=p
}
###############################

#filter
p=p[p$PlotColumn%in%'ncRNA','Same_ncRNA',drop=F]

#if there are remaining types
if (nrow(p)>0) {
  if(length(unique(p$Same_ncRNA)[unique(p$Same_ncRNA)%in%names(nc_color)==F])>0){
    print(paste0('unexpected ncRNA category: ',unique(p$Same_ncRNA)[unique(p$Same_ncRNA)%in%names(nc_color)==F]))
  }else{
    nc_color_sel=nc_color[unique(p$Same_ncRNA)]
    nc_color_sel=nc_color_sel[is.na(nc_color_sel)==F]  
    p$RNA=""
    gg=ggplot(p,aes(fill=Same_ncRNA,x=RNA) ) +
        geom_bar(width=.5) +
        theme_classic() +
        ggtitle("ncRNA only: # of peaks per category") +
        labs(fill='RNA Type') +
        scale_color_manual(values=nc_color_sel) +
        scale_fill_manual(values=nc_color_sel) + 
        blank_theme + 
        coord_polar("y",start=0)
    (gg) ###updating print to be consistent
    #ggplotly(gg)
  }  
}else{print("No ncRNA")}
```

<br>  
<br> 

**III.2 Peak counts do not include rRNA peaks**   


```{r PeaksPerCategory_NO_rRNA, fig.asp=.5, fig.align='center',fig.height=4, fig.width=8,echo=F,warning=F,eval=T,include=T}
p=Peaks_Anno
p$PlotColumn=p$Same_Comb_type
p=p[!p$Same_Comb_type%in%'rRNA',]
if (nrow(p)>0) {
p[is.na(p$Same_Comb_type),'PlotColumn']='no Feature'
colnames(p)[colnames(p)%in%'PlotColumn']='RNA_Type'
p$RNA=''
#plot
gg=ggplot(p,aes(fill=RNA_Type,x=RNA) )+
  geom_bar(position="stack",width=.5) +
  theme_classic() +
  ggtitle("BioType: # of peaks per category") +
  scale_fill_manual(values=class_color) 
ggplotly(gg)
}else{print("No ncRNA")}
```


```{r PeaksPer_NCRNA_Category_NO_rRNA, fig.asp=.5, fig.align='center',fig.height=4, fig.width=8,echo=F,warning=F,eval=T,include=T}
p=Peaks_Anno
p$PlotColumn=p$Same_Comb_type
p=p[!p$Same_Comb_type%in%'rRNA',]
p=p[p$PlotColumn%in%'ncRNA','Same_ncRNA',drop=F]
#if there are remaining types
if (nrow(p)>0) {
  if(length(unique(p$Same_ncRNA)[unique(p$Same_ncRNA)%in%names(nc_color)==F])>0){
    print(paste0('unexpected ncRNA category: ',unique(p$Same_ncRNA)[unique(p$Same_ncRNA)%in%names(nc_color)==F]))
}else{
  nc_color_sel=nc_color[unique(p$Same_ncRNA)]
  nc_color_sel=nc_color_sel[is.na(nc_color_sel)==F]  
    p$RNA=""
    gg=ggplot(p,aes(fill=Same_ncRNA,x=RNA) ) +
      geom_bar(width=.5)+theme_classic() +
      ggtitle("ncRNA only: # of peaks per category") +
      labs(fill='ncRNA Type') +
      scale_color_manual(values=nc_color_sel) + 
      scale_fill_manual(values=nc_color_sel) +
      blank_theme + 
      coord_polar("y",start=0)
    (gg)
    # ggplotly(gg)
}
}else{print("No ncRNA")}
```

**III.3 Peak counts do not include rRNA peaks nor peaks without unique reads**   


```{r PeaksPerCategory_NO_rRNA_unique, fig.asp=.5, fig.align='center',fig.height=4, fig.width=8,echo=F,warning=F,eval=T,include=T}
p=Peaks_Anno
p$PlotColumn=p$Same_Comb_type
p=p[!p$Same_Comb_type%in%'rRNA',]
p=p[p$Counts_Unique > 0,]
if (nrow(p)>0) {
p[is.na(p$Same_Comb_type),'PlotColumn']='no Feature'
colnames(p)[colnames(p)%in%'PlotColumn']='RNA_Type'
p$RNA=''
#plot
gg=ggplot(p,aes(fill=RNA_Type,x=RNA) )+
  geom_bar(position="stack",width=.5) +
  theme_classic() +
  ggtitle("BioType: # of peaks per category") +
  scale_fill_manual(values=class_color) 
ggplotly(gg)
}else{print("No ncRNA")}
```


```{r PeaksPer_NCRNA_Category_NO_rRNA_unique, fig.asp=.5, fig.align='center',fig.height=4, fig.width=8,echo=F,warning=F,eval=T,include=T}
p=Peaks_Anno
p$PlotColumn=p$Same_Comb_type
p=p[!p$Same_Comb_type%in%'rRNA',]
p=p[p$Counts_Unique > 0,]
p=p[p$PlotColumn%in%'ncRNA','Same_ncRNA',drop=F]
#if there are remaining types
if (nrow(p)>0) {
  if(length(unique(p$Same_ncRNA)[unique(p$Same_ncRNA)%in%names(nc_color)==F])>0){
    print(paste0('unexpected ncRNA category: ',unique(p$Same_ncRNA)[unique(p$Same_ncRNA)%in%names(nc_color)==F]))
}else{
  nc_color_sel=nc_color[unique(p$Same_ncRNA)]
  nc_color_sel=nc_color_sel[is.na(nc_color_sel)==F]  
    p$RNA=""
    gg=ggplot(p,aes(fill=Same_ncRNA,x=RNA) ) +
      geom_bar(width=.5)+theme_classic() +
      ggtitle("ncRNA only: # of peaks per category") +
      labs(fill='ncRNA Type') +
      scale_color_manual(values=nc_color_sel) + 
      scale_fill_manual(values=nc_color_sel) +
      blank_theme + 
      coord_polar("y",start=0)
    (gg)
    # ggplotly(gg)
}
}else{print("No ncRNA")}
```    

<br>


### VIII. Number of reads assigned to each category   

#### VIII.1 Counts for RNA biotype  


```{r ReadsPerCategory_1, fig.asp=.5, fig.align='center',fig.height=4, fig.width=8,echo=F,warning=F,include=F,eval=T}

p=Peaks_Anno
p$PlotColumn=p$Same_Comb_type
p$PlotColumn=as.character(p$PlotColumn)
#p[p$Same_Repeat%in%'rRNA','PlotColumn']='rRNA'
p[p$Same_ncRNA%in%'yRNA','PlotColumn']='yRNA'
p$PlotColumn=factor(p$PlotColumn, levels = c("ncRNA",
                                              "rRNA",
                                              "yRNA",
                                              "protein_coding: exon",
                                              "Repeat Element",
                                              "pseudogene",
                                              "protein_coding: Intron",
                                              "lncRNA",
                                              "no Feature"))#,
                                              #"Antisense Feature"))

u=c("ncRNA",
     "rRNA",
     "yRNA",
     "protein_coding: exon",
     "Repeat Element",
     "pseudogene",
     "protein_coding: Intron",
     "lncRNA",
     "no Feature")#,
     #"Antisense Feature",)

if (nrow(p)>0) {
fcount=data.frame(u,1:length(u));colnames(fcount)=c('RNA_Type','order');fcount$Counts_Unique=NA;fcount$Counts_MM=NA;#fcount$Peak_count=NA

for (x in 1:length(u)) {
  pam=p[p$PlotColumn%in%u[x],]
  fcount[x,1]=u[x]
  fcount[x,'Counts_Unique']=colSums(pam[,'Counts_Unique',drop=F],na.rm = T)
  fcount[x,'Counts_fracMM']=colSums(pam[,'Counts_fracMM',drop=F],na.rm = T)
    fcount[x,'Peak_count']=nrow(pam)
fcount$RNA=""
fcount$RNA_Type=factor(fcount$RNA_Type,
  levels = c("ncRNA",
             "rRNA",
             "yRNA",
             "protein_coding: exon",
             "Repeat Element",
             "pseudogene",
             "protein_coding: Intron",
             "lncRNA",
             "no Feature"))#,
             #"Antisense Feature"))
}
} else {print('No Features')}

```{r ReadsPerCategory_2, fig.asp=.5, fig.align='center',fig.height=4, fig.width=8,echo=F,warning=F,include=T,eval=T}

if (exists('fcount')) {
  gg1 = ggplot(fcount,
               aes(fill=RNA_Type,
                   x=RNA,
                   y=Counts_Unique)) +
        geom_bar(stat='identity',
                 position="stack",
                 width=.5) +
                 theme_classic() +
        ggtitle("BioType - # of reads (unique) by category)") +
        scale_fill_manual(values=class_color)  
  ggplotly(gg1)
} else {print('No Features')}
```

```{r ReadsPerCategory_3, fig.asp=.5, fig.align='center',fig.height=4, fig.width=8,echo=F,warning=F,include=T,eval=T}
if (exists('fcount')) {
gg2 = ggplot(fcount,aes(fill=RNA_Type,x=RNA,y=Counts_fracMM)) +
  geom_bar(stat='identity',position="stack",width=.5) +
  theme_classic() +
  ggtitle("BioType - # of reads (unique + Frac. MM) by category)") +
  scale_fill_manual(values=class_color)
ggplotly(gg2)
} else {print('No Features')}
```
<br>



#### VIII.2 Counts for ncRNA      


```{r ReadsPerCategory_ncRNA_1, fig.asp=.5, fig.align='center',fig.height=4, fig.width=8,echo=F,warning=F,include=F,eval=T}
p=Peaks_Anno
p$PlotColumn=p$Same_Comb_type
p$PlotColumn=as.character(p$PlotColumn)
#p[p$Same_Repeat%in%'rRNA','PlotColumn']='rRNA'
p[p$Same_ncRNA%in%'yRNA','PlotColumn']='yRNA'
p$PlotColumn=factor(p$PlotColumn, levels = c("ncRNA",
                                             "rRNA",
                                             "yRNA",
                                             "protein_coding: exon",
                                             "Repeat Element","pseudogene",
                                             "protein_coding: Intron",
                                             "lncRNA",
                                             "no Feature"))#,
                                             #"Antisense Feature"))

p=p[p$PlotColumn%in%'ncRNA',]
if (nrow(p)>0) {
  u=unique(p$Same_ncRNA)
  nccount=data.frame(u,1:length(u));colnames(nccount)=c('Same_RNA_Subtype','Counts_All');nccount$Counts_Unique=NA;#nccount$Counts_MM=NA;nccount$Peak_count=NA
  
  for (x in 1:length(u)) {
    pam=p[p$Same_ncRNA%in%u[x],]
    nccount[x,1]=u[x]
    nccount[x,'Counts_Unique']=colSums(pam[,'Counts_Unique',drop=F],na.rm = T)
    nccount[x,'Counts_fracMM']=colSums(pam[,'Counts_fracMM',drop=F],na.rm = T)
    nccount[x,'Peak_count']=nrow(pam)
  }
}else{print('No Significant ncRNA CLIP peaks')}
```

```{r ReadsPerCategory_ncRNA_2, fig.asp=.5, fig.align='center',fig.height=4, fig.width=8,echo=F,warning=F,include=T,eval=T}
if (exists('nccount')) {
nccount$RNA=''
if(length(unique(nccount$Same_RNA_Subtype)[unique(nccount$Same_RNA_Subtype)%in%names(nc_color)==F])>0){
  print(paste0('unexpected ncRNA category: ',unique(nccount$Same_RNA_Subtype)[unique(nccount$Same_RNA_Subtype)%in%names(nc_color)==F]))
}else{
nc_color_sel=nc_color[unique(nccount$Same_RNA_Subtype)]
nc_color_sel=nc_color_sel[is.na(nc_color_sel)==F]
#plot
gg1 = ggplot(nccount,aes(fill=Same_RNA_Subtype,x=RNA,y=Counts_Unique) ) +
  geom_bar(stat='identity',width=.5)+theme_classic() +
  ggtitle("ncRNA only: # of reads (unique) by category") + 
  labs(color='RNA Type') +
  scale_color_manual(values=nc_color_sel)+scale_fill_manual(values=nc_color_sel) + 
  blank_theme + 
  coord_polar("y",start=0)
plot(gg1)
}
}else{print('No Unique counts for ncRNA peaks')}
#plot
```
```{r ReadsPerCategory_ncRNA_3, fig.asp=.5, fig.align='center',fig.height=4, fig.width=8,echo=F,warning=F,include=T,eval=T}
  
if (exists('nccount')) {
nccount$RNA=''
if(length(unique(nccount$Same_RNA_Subtype)[unique(nccount$Same_RNA_Subtype)%in%names(nc_color)==F])>0){
  print(paste0('unexpected ncRNA category: ',unique(nccount$Same_RNA_Subtype)[unique(nccount$Same_RNA_Subtype)%in%names(nc_color)==F]))
}else{
nc_color_sel=nc_color[unique(nccount$Same_RNA_Subtype)]
nc_color_sel=nc_color_sel[is.na(nc_color_sel)==F]
gg2 = ggplot(nccount,aes(fill=Same_RNA_Subtype,x=RNA,y=Counts_fracMM) ) +
  geom_bar(stat='identity',width=.5)+theme_classic() +
  ggtitle("ncRNA only: # of reads (unique + Frac. MM) by category") + 
  labs(color='RNA Type') +
  scale_color_manual(values=nc_color_sel)+scale_fill_manual(values=nc_color_sel) +
  blank_theme + 
  coord_polar("y",start=0)
plot(gg2)
}
}else{print('No unique + Frac. MM counts for ncRNA CLIP peaks')}
```      

<br>



### IV. Read Counts by peak for each category    
  
#### IV.1 Unique Read counts (Count Unique reads only)   

<br>   

```{r fig.asp=.5, fig.align='center',fig.height=6, fig.width=8,echo=F,warning=F,eval=T, include=T}
p=Peaks_Anno
p$PlotColumn=p$Same_Comb_type
if (incd_rRNA==FALSE) {
  p=p[!p$Same_Comb_type%in%'rRNA',]
}
colnames(p)[colnames(p)%in%'PlotColumn']='RNA_Type'
if (nrow(p)>0) {
p$RNA=''
plot
gg = ggplot(p,aes(x=RNA_Type,y=Counts_Unique,fill=RNA_Type)) +
            geom_violin() +
            theme_classic() +
            geom_jitter(height = 0, width = 0.1,size=0.5) +
            scale_y_continuous(trans = "log2", oob = squish_infinite) +
            ggtitle('Peak Counts (Unique) for\nby Feature') +
            ylab("Counts") +
            xlab("") +
            theme(plot.title = element_text(hjust = 0.5,size = 15),
                               axis.title=element_text(size=15),
                               axis.text=element_text(size=12)) +
            theme(legend.position = "none") +
            theme(legend.position = "none") +
            theme(axis.text.x = element_text(angle = -45))+
            scale_fill_manual(values=class_color) 
if (nrow(p)==1|nrow(p)>5000) {gg
  }else { ggplotly(gg)}
}else{print('No Features')}
```  
<br/>    


```{r fig.asp=.5, fig.align='center',fig.height=6, fig.width=8,echo=F,warning=F,eval=T, include=T}
p=Peaks_Anno
p$PlotColumn=p$Same_ncRNA
  p=p[p$Same_Comb_type%in%'ncRNA',]
colnames(p)[colnames(p)%in%'PlotColumn']='RNA_Type'
if (nrow(p)>0) {
  p$ID=paste0(p$PeakID,"\n",p$Same_gene_name_comb)
  
  if(length(unique(p$RNA_Type)[unique(p$RNA_Type)%in%names(nc_color)==F])>0){
  print(paste0('unexpected ncRNA category: ',unique(p$RNA_Type)[unique(p$RNA_Type)%in%names(nc_color)==F]))
}else{
  nc_color_sel=nc_color[unique(p$RNA_Type)]
  nc_color_sel=nc_color_sel[is.na(nc_color_sel)==F]
plot
gg = ggplot(p, aes(x=RNA_Type,y=Counts_Unique,fill=RNA_Type)) +
      geom_violin() +
      theme_classic() +
      geom_jitter(height = 0, width = 0.1,size=0.5) +
      scale_y_continuous(trans = "log2", oob = squish_infinite) +
      ggtitle('Peak Counts (Unique) for\nby ncRNA Type') +
      ylab("Counts") +
      xlab("") +
      theme(plot.title = element_text(hjust = 0.5,size = 15),
            axis.title = element_text(size=15),
            axis.text = element_text(size=12)) +
      theme(legend.position = "none") +
      theme(legend.position = "none") +
      theme(axis.text.x = element_text(angle = -45)) +
      scale_fill_manual(values=nc_color_sel)
if (nrow(p)==1|nrow(p)>5000) {gg
}else { ggplotly(gg,tooltip = c("text", "y"))}
}
}else{print('No Features')}
```
<br>   

#### IV.2 Fractional Multimap Read counts \n (Count Unique reads + (MultiMapped read / # of different mappings) )     

Does include rRNA
<br>   

```{r fig.asp=.5, fig.align='center',fig.height=6, fig.width=8,echo=F,warning=F,eval=T, include=T}
p=Peaks_Anno
p$PlotColumn=p$Same_Comb_type
colnames(p)[colnames(p)%in%'PlotColumn']='RNA_Type'
if (nrow(p)>0) {
p$RNA=''
#plot
gg=ggplot(p,aes(x=RNA_Type,y=Counts_fracMM,fill=RNA_Type)) +
    geom_violin() +
    theme_classic() +
    geom_jitter(height = 0, width = 0.1,size=0.5) +
    scale_y_continuous(trans = "log2", oob = squish_infinite) +
    ggtitle('Peak Counts (Unique + Frac. MM)\nby Feature') +
    ylab("Counts") +
    xlab("") +
    theme(plot.title = element_text(hjust = 0.5,size = 15),
          axis.title = element_text(size=15),
          axis.text = element_text(size=12)) +
    theme(legend.position = "none") +
    theme(axis.text.x = element_text(angle = -45)) +
    scale_fill_manual(values=class_color) 
if (nrow(p)==1|nrow(p)>5000) {gg
  }else { ggplotly(gg)}
}else {print("No Features")}
```  
<br/>    

<br/>    

```{r fig.asp=.5,fig.align='center',fig.height=6, fig.width=8,echo=F,warning=F,eval=T}
p=Peaks_Anno
p$PlotColumn=p$Same_ncRNA
  p=p[p$Same_Comb_type%in%'ncRNA',]
colnames(p)[colnames(p)%in%'PlotColumn']='RNA_Type'
if (nrow(p)>0) {
p$ID=paste0(p$PeakID,"\n",p$Same_gene_name_comb)
  
if(length(unique(p$RNA_Type)[unique(p$RNA_Type)%in%names(nc_color)==F])>0){
  print(paste0('unexpected ncRNA category: ',unique(p$RNA_Type)[unique(p$RNA_Type)%in%names(nc_color)==F]))
}else{  
  nc_color_sel=nc_color[unique(p$RNA_Type)]
  nc_color_sel=nc_color_sel[is.na(nc_color_sel)==F]
#plot
  gg=ggplot(p,aes(x=RNA_Type,y=Counts_fracMM,fill=RNA_Type)) +
      geom_violin() +
      theme_classic() +
      geom_jitter(height = 0, width = 0.05,size=.5) +
      scale_y_continuous(trans = "log2", oob = squish_infinite) +
      ggtitle('Peak Counts (Unique + Frac. MM)\nby ncRNA Type') +
      ylab("Counts") +
      xlab("") +
      theme(plot.title = element_text(hjust = 0.5,size = 15),
            axis.title = element_text(size=15),
            axis.text=element_text(size=12)) +
      theme(legend.position = "none") +
      theme(axis.text.x = element_text(angle = -45)) +
      scale_fill_manual(values=nc_color_sel) #+ylim(min(p$Counts_Unique),max(p$Counts_Unique))
  if (nrow(p)==1|nrow(p)>5000) {gg
  }else { ggplotly(gg,tooltip = c("text", "y"))}
}
}else {print("No Features")}
```  
<br>
<br>  


### V. Opposite strand CLIP peak attributes   


```{r , fig.align='center',fig.height=6, fig.width=12,echo=F,warning=F}
p=Peaks_Anno
#p=p$Oppo_Comb_type
###phil need else statement
if (incd_rRNA==FALSE) {
  p=p[!p$Oppo_Comb_type%in%'rRNA',]
}
if (nrow(p)>0) {
colnames(p)[colnames(p)%in%'Oppo_Comb_type']='RNA_Type'
p$RNA=''
#plot
gg = ggplot(p,aes(fill=RNA_Type,x=RNA_Type)) +
  geom_bar(stat="count") +
  theme_classic() +
  ggtitle("Opposite Strand Feature - Number of peaks by category") +
  xlab("") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title = element_text(size=15),
        axis.text = element_text(size=10),
        axis.text.x = element_text(angle = -45,vjust = .5)) +
  scale_fill_manual(values=class_color) 
ggplotly(gg)
}else {print("No Features")}
```


```{r fig.asp=.5, fig.align='center',fig.height=6, fig.width=12,echo=F,warning=F,eval=T,incude=T}

print('SKIP')

#p=Peaks_Anno
##phil need else statement
#if (incd_rRNA==FALSE) {
#  p=p[!p$Oppo_Comb_type%in%'rRNA',]
#}
##p=p[p$Comb_type_exon_Oppo%in%'Antisense Feature',]
#if (nrow(p)>0) {
#  p$transition=paste0(p$Same_Comb_type,'->\n',p$Oppo_Comb_type)
#colnames(p)[colnames(p)%in%'transition']='RNA_Type'
#p$Difference=''
##plot
#gg = ggplot(p,aes(fill=Same_Comb_type,x=factor(Oppo_Comb_type))) +
#      geom_bar(stat="count") +
#      theme_classic() +
#      ggtitle("Opposite Strand Feature Features: \n Compare Antisense and Sense stand Annotation") +
#      xlab("Antisense strand Annotation") +
#      ylab("Peak Count") +
#      labs(fill = "Sense Stand Annotation") +
#      theme(plot.title = element_text(hjust = 0.5,size=20),
#            axis.title = element_text(size=15),
#            axis.text = element_text(size=10),
#            legend.position = "right",
#            axis.text.x = element_text(angle = -45,vjust = .5))
#ggplotly(gg)
#}else {print("No Features")}
```

<br>  
